Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-11-13 xrb2/paper.tex
Reorganized subsections to be rotation & temperature-specific, added some temperature discussion

In our latest set of runs, we explore a variety of initial rotation rate and temperature conditions for the flame. The initial conditions are given in Table \ref{table:500Hz_runs}.

@@

\subsection{Flame Propagation}

@@

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0$.}
	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
	\startdata
	{\tt flame\_wave\_500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
	{\tt flame\_wave\_1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
	{\tt flame\_wave\_500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
	{\tt flame\_wave\_500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
	\enddata
	%
\end{deluxetable}

Figures~\ref{fig:flame_speeds_1} and \ref{fig:flame_speeds_2} give position vs.\ time for our runs, with the former showing dependence
on rotation rate and the latter on the neutron star temperature. As some of the runs (in particular the $\tt{flame\_wave\_3E8}$ run) exhibit
significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit
parameters. These results are given in Table~\ref{table:flame_speeds}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
and reach higher burning rates, resulting in an $\epsdot$ increase over time.

@@

{\color{red} Maybe Nicole can talk a bit about the effect of the neutron star temperature here? I have a placeholder for the plot earlier, but
feel free to move/edit it. What's in this section now is just an initial draft/outline.}

@@

\subsection{Effect of Temperature Structure}

\begin{figure}[t]
\centering
\plotone{compare_500Hz_abar}
\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope
structures.  We see that the hottest model has a faster propagating flame.}
\end{figure}

Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.

>>>>

In our latest set of runs, we explore a variety of initial rotation rate and temperature conditions for the flame. The initial conditions are provided in Table \ref{table:500Hz_runs}.

\subsection{Effect of Rotation Rate on Flame Structure}\label{ssec:rot_structure}

@@

\subsection{Effect of Rotation Rate on Flame Propagation}\label{ssec:rot_propagation}

@@

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0$.}
	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
	\startdata
	{\tt flame\_wave\_1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
	{\tt flame\_wave\_500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
	{\tt flame\_wave\_500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
	{\tt flame\_wave\_500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
	\enddata
	%
\end{deluxetable}

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{flame\_wave\_500}$ and $\tt{flame\_wave\_1000}$ runs to show the dependence on rotation rate. As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters. These results are given in Table~\ref{table:flame_speeds}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
and reach higher burning rates, resulting in an $\epsdot$ increase over time.

@@

@@

\subsection{Effect of Temperature Structure}\label{ssec:temp_structure}

To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:500Hz_runs}. For all the 500 Hz simulations (with the exception of the coolest run, $\tt{flame\_wave\_500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$. If we let $T_\mathrm{star} < T_\mathrm{hi}$, the cooler neutron star surface might act as a heat sink and siphon away energy that would otherwise go into heating the burning layer.

\begin{figure}[t]
\centering
\plotone{compare_500Hz_abar}
\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope structures. } %This will be updated to show at 70ms
\end{figure}

Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures. The hottest model, $\tt{flame\_wave\_500\_3E8}$ has a faster propagating flame. It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm). It is likely that a small amount of hot material broke through the top of the atmosphere.

{\color{red} On the subject of things breaking out of the atmosphere, discuss interesting characteristics of the extreme 4E8 run here: maybe include a plot showing abar and temp at 40 ms, and another plot showing the momentum components?}

{\color{red} Discuss flame acceleration here, it's sort of mysterious but could be due to energy released from burning dominating the flame propagation and increasing speed over time, or maybe enhanced mixing/diffusion is pulling in more fuel to burn, or maybe the average fuel layer temp is going up, making it easier for the flame to burn and propagate}
